---
title: "Reproducible Research: Peer Assessment 1"
author: "Anizah Nordin"
date : "13 October 2015"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Steps as follow:

1.  Load the data (i.e. read.csv())

```{r}
dActivity <- read.csv("activity.csv")
```

2.  Check if data need to be transform into a suitable format for analysis

```{r}
str(dActivity)

```

## What is mean total number of steps taken per day?

1.  Make a histogram of the total number of steps taken each day (ignore missing values)

```{r}
totalSteps <- aggregate(steps ~ date, dActivity, sum)
barplot(totalSteps$steps, names.arg = totalSteps$date, xlab = "Date", ylab = "Steps", col="blue", main="Total Steps Taken Per Day")
```

2.  Calculate and report the mean and median total number of steps taken per day (remove missing values)

```{r}
mean(totalSteps$steps, na.rm = TRUE)
median(totalSteps$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?

1.  Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
stepsInterval <- aggregate(steps ~ interval, dActivity, mean)
plot(stepsInterval, type = "l", main="Average Daily Activity Pattern", xlab="Interval", ylab="Average Number of Steps", col="blue")

```

2.  Which 5-minute interval, on average across all the days in the dataset,   
contains the maximum number of steps?

```{r}
stepsInterval$interval[which.max(stepsInterval$steps)]
```

## Imputing missing values

1.  Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)  

```{r}
sum(is.na(dActivity))
```

2.  Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
dActivity <- merge(dActivity, stepsInterval, by = "interval", suffixes = c("", 
    ".y"))
nas <- is.na(dActivity$steps)
dActivity$steps[nas] <- dActivity$steps.y[nas]
dActivity <- dActivity[, c(1:3)]

```

3.  Make a histogram of the total number of steps taken each day.   

```{r}
stepsDay <- aggregate(steps ~ date, dActivity, sum)
barplot(stepsDay$steps, names.arg = stepsDay$date, xlab = "Date", ylab = "Steps taken each day")


```

4.  Calculate and report the mean and median total number of steps taken per day. 
```{r}
mean(stepsDay$steps)
median(stepsDay$steps)

```

## Are there differences in activity patterns between weekdays and weekends?

1.  Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
dayType <- function(date) {
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        "Weekend"
    } else {
        "Weekday"
    }
}
dActivity$dayType <- as.factor(sapply(dActivity$date, dayType))

```

2.  Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
par(mfrow = c(2, 1))
for (type in c("Weekday", "Weekend")) {
    stepsType <- aggregate(steps ~ interval, data = dActivity, subset = dActivity$dayType == type, mean)
    plot(stepsType, type = "l", main = type, col="blue", lwd = 2) 
    
}
```

